---
title: "Publishing the art"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(arttools)
```

This article still needs to be written. For what it's worth, the process is essentially this:

1. Move the to-be-published folder from `repo_local_path()` into `bucket_local_path()`
2. Upload the folder from the local bucket to the remote bucket (using `gcloud` CLI for now, because that's what I use)
3. Add the relevant page to your website.

Yes, I know. This is totally a "draw the rest of the owl" kind of thing. But I haven't gotten to the point where I've systematised this part of the workflow, and so for the time being this page is basically a promissory note!

<!--
### Uploading curated images

- Optional, but desirable: move (or copy) the gallery folder to the `~/Bucket/djnavarro-art` folder. Gallery folders properly belong to the object store rather than the git repo, so we might as well have that reflected locally!
- To publish a gallery folder, copy it to the `djnavarro-art` bucket

  ```
  gcloud storage cp series-fake gs://djnavarro-art --recursive --project generative-art-389407
  ```
  
### Creating a webpage

-->
