---
title: "Creating art repositories"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  arttools.bucket.remote = "https://storage.googleapis.com/djnavarro-art",
  arttools.repos.remote = "https://github.com/djnavarro"
)
```

```{r setup}
library(arttools)
options(
  arttools.repos.local = file.path(tempdir(), "temp_repos"),
  arttools.bucket.local = file.path(tempdir(), "temp_bucket")
)
```

### Repository name

One of the first choices you have to make when setting up a new system is to choose a name for the repository. The arttools package doesn't enforce any naming system, but from my personal experience I've learned that it's useful to adopt a consistent `[prefix]-[system name]` format. All my art repositories use "series" as a prefix, which makes it a lot easier for me to find them on GitHub later! For example, if my generative art system is called "cards", the repository name would be "series-cards". Any time a function in the arttools package takes a `series` argument, I'd write `series = "series-cards"` or something along those lines.

There's nothing special about using "series" as the prefix though. I chose that solely because of how I think about generative art systems: to me they're tools for creating an "art series", so that seems to make sense to me.

### Repository folder structure

You can use `repo_create()` to create the skeleton for a local series repository. If my generative art system is called "fake", my series name will be `"series-fake"`. I might use a command like this to create a skeleton:

```{r}
repo_create(series = "series-fake", license = "ccby")
```

This is the file tree it creates:

```{r}
#| echo: false
fs::dir_tree(
  repo_local_path("series-fake"), 
  all = TRUE, 
  recurse = TRUE
)
```

Note that:

- It writes a `.gitignore` file to ensure git won't commit output files
- It writes a `.here` file to ensure `here::here()` will always work

But also note that:

- It doesn't create an RStudio project: use `usethis::create_project()` for that
- It doesn't initialise a git repository: use `usethis::use_git()` for that
- It doesn't create a remote on GitHub: use `usethis::use_github()` for that

The other two top-level files it writes are:

- `README.md` is always created
- `LICENSE.md` is created only if `license` is specified

There four folders in the repository:

- `source` is for source code for the generative art system
- `input` is for other input files used by the system (optional)
- `build` is for ancillary scripts that are not part of the system (optional)
- `output` is for output files generated by the system (ignored by git)

Within the source folder there are three other files:

- `source/common.R`
- `source/art-system_01.R`
- `source/art-system_02.R`

These files define a simple generative art system that I'll use to discuss other aspects of generative art workflow in another article.

```{r}
#| echo: false
#| results: hide
fs::dir_delete(repo_local_path("series-fake"))
```


