---
title: "Creating art repositories"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(
  arttools.bucket.remote = "https://storage.googleapis.com/djnavarro-art",
  arttools.bucket.local = "~/Buckets/djnavarro-art",
  arttools.repos.remote = "https://github.com/djnavarro",
  arttools.repos.local = "~/GitHub"
)
```

```{r setup}
library(arttools)
```


### Repository name

One of the first choices you have to make when setting up a new system is to choose a name for the repository. The arttools package doesn't enforce any naming system, but from my personal experience I've learned that it's useful to adopt a consistent `[prefix]-[system name]` format. All my art repositories use "series" as a prefix, which makes it a lot easier for me to find them on GitHub later! For example, if my generative art system is called "cards", the repository name would be "series-cards". Any time a function in the arttools package takes a `series` argument, I'd write `series = "series-cards"` or something along those lines.

There's nothing special about using "series" as the prefix though. I chose that solely because of how I think about generative art systems: to me they're tools for creating an "art series", so that seems to make sense to me.

### Repository folder structure

You can use `repo_create()` to create the skeleton for a local series repository. If my generative art system is called "fake", my series name will be `"series-fake"`. I might use a command like this to create a skeleton:

```{r}
repo_create(
  series = "series-fake",
  license = "ccby",
  destination = tempdir()
)
```

Note that the `destination = tempdir()` part is purely for the sake of this article. By default, the `repo_create()` function assumes you want to create the skeleton in the same local folder that you have all your other art repositories, and as such the default value of `destination = repo_local_path()` is probably what you actually want. This is the file tree it creates:

```{r}
#| echo: false
series_path <- fs::path(tempdir(), "series-fake")
fs::dir_tree(
  series_path, 
  all = TRUE, 
  recurse = TRUE
)
```

```{r}
#| echo: false
#| results: hide
fs::dir_delete(series_path)
```

Note that:

- It writes a `.gitignore` file to ensure git won't commit output files
- It writes a `.here` file to ensure `here::here()` will always work

But also note that:

- It doesn't create an RStudio project: use `usethis::create_project()` for that
- It doesn't initialise a git repository: use `usethis::use_git()` for that
- It doesn't create a remote on GitHub: use `usethis::use_github()` for that

The other two top-level files it writes are:

- `README.md` is always created
- `LICENSE.md` is created only if `license` is specified

There are three folders that will be tracked by git:

- `source` should contain source code for the generative art system
- `input` should contain other input files used by the system (optional)
- `build` should contain ancillary scripts that are not part of the art system (optional)

In practice I find that I don't always need the `input` and `build` folders and often delete them. The other two folders are used to store image files. These are:

- `output` is for raw output files generated directly by the system
- `series-fake` is for the manually curated versions of the output files, and is the folder that you'd later move into cloud storage once you're done with the system

Finally, notice that within the source folder there are three other files:

- `source/common.R`
- `source/art-system_01.R`
- `source/art-system_02.R`

These files define a simple generative art system that I'll use to discuss other aspects of generative art workflow.
